---
title: "Logistic Regression"
author: "Linda"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Load the libraries

```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(caret)
library(fastDummies)
```

## Load the dataset


```{r}
data<-read.xlsx(file.choose())
```

## Exploratory Data Analysis

Explore the dataset

```{r, echo=TRUE,eval=TRUE}
head(data) # 6 first rows

```
```{r}
summary(data) #Descriptive statistics
```



 Data cleaning
```{r}
sum(is.na(data))
##Check missing values
data$last_review[is.na(data$last_review)]<-0
data$name[is.na(data$name)]<-'Unknown'
data$host_name[is.na(data$host_name)]<-'Unkown'
data$reviews_per_month[is.na(data$reviews_per_month)]<-mode(data$reviews_per_month)[1]
```
```{r}
## Check the absence of na values
colSums(is.na(data))
```
```{r}
###Check for outliers in the numerical variables using boxplots
data|>ggplot(aes(price))+geom_boxplot()+theme_bw()

```
```{r}
data|>ggplot(aes(minimum_nights))+geom_boxplot()+theme_bw()
```
```{r}
data|>ggplot(aes(number_of_reviews))+geom_boxplot()+theme_bw()
```
```{r}
data|>ggplot(aes(availability_365))+geom_boxplot()+theme_bw()
```
```{r}
data|>ggplot(aes(calculated_host_listings_count))+geom_boxplot()+theme_bw()
```
```{r}
##Removing outliers
data<-data|>group_by(room_type)|>
  mutate(Q1=quantile(price,.25,na.rm=TRUE),
         Q3=quantile(price,.75,na.rm=TRUE),
         IQR=Q3-Q1,
         Lower_bound=Q1 -1.5*IQR,
         Upper_bound=Q3+ 1.5*IQR)|>
  filter(price>=Lower_bound & price <= Upper_bound)|>
  ungroup()|>
  select(-Q1,-Q3,-IQR)
summary(data$price)
```
```{r}
data<-data|>
  mutate(Q1=quantile(minimum_nights,0.25,na.rm=TRUE),
                    Q3=quantile(minimum_nights,0.75,na.rm=TRUE),
                    IQR=Q3-Q1,
                    Lower_bound=Q1-1.5*IQR,
                    Upper_bound=Q3+1.5*IQR)|>
  filter(minimum_nights>=Lower_bound & minimum_nights<=Upper_bound)|>
  select(-Q1,-Q3,-IQR)
summary(data$minimum_nights)
```




## Detect and Remove the duplicated rows 
```{r}
sum(duplicated(data))
data<-data|> ###Removes the duplicated rows
  distinct()

```


## Select the columns to use to determine what happens when the price of a listing is higher than the mean price of the listings per night using the minimum nights requirement,availability of a listing during the year, the number of reviews and room type.
```{r}
data<-data|>
  select('price','minimum_nights','number_of_reviews','availability_365','room_type')
head(data) ## 6 first rows
```


Histograms for numerical variables
```{r}
data|>ggplot(aes(price))+
  geom_histogram(binwidth = 20,position ='identity')+
  theme_bw()+
  labs(x='Price',y='Frequencies',title='Distribution of price per night')
```
```{r}
data|>ggplot(aes(minimum_nights))+geom_histogram(binwidth = 2)+
  theme_bw()+
  labs(x='Minimum nights',y='Frequencies',title='Distribution of the minimum nights')
```
```{r}
data|>ggplot(aes(availability_365))+geom_histogram(binwidth = 50 )+
  theme_bw()+
  labs(x='Availability',y='Frequencies',title='Distribution of the availability ')
```
## Bar chart for the room type proportions
```{r}
data|>ggplot(aes(room_type,fill=room_type))+geom_bar()+
  theme_bw()+labs(x='Room type',y='Frequencies',title='Pie chart room type proportion')


```

## Converting categorical variables to numerical variables
```{r}
data$price_class<-ifelse(data$price>mean(data$price,na.rm=TRUE),1,0)  #Add column for converted price values
##data<-dummy_cols(data, select_columns ='room_type' )##Create room type dummy variables
head(data)
```

## Data partitioning
Splitting the dataset into train and test
```{r}
set.seed(1124)#For reproducibility
library(caret)
index<-createDataPartition(data$price_class,
        p=.7,list=FALSE)
data_train<-data[index,]
data_test<-data[-index,]
head(data_train)
```

## Build the model
This involves training the model on the expected output
```{r}
model<-glm(price_class~minimum_nights+number_of_reviews+availability_365+room_type,family=binomial,data=data_train)
summary(model)
```

## Test the model
```{r}
data_test$predicted_probabilities<-predict(model,newdata=data_test,type='response')
##Convert the predicted probabilities to binary of 1's and 0's with a threshold of .5
data_test$predicted_class<-ifelse(data_test$predicted_probabilities>.5,1,0)

head(data_test)

```


## Validate the model
```{r}
table(prediction=data_test$predicted_class, Actual=data_test$price_class)
```

## Confusion matrix 
To validate the model accuracy and precision
```{r}
confusion_Matrix<-confusionMatrix(table(prediction=data_test$predicted_class, Actual=data_test$price_class))
###Get the confusion matrix
confusion_Matrix
```

##Histogram 
To visualize the predicted probanilities
```{r}
data_test|>ggplot(aes(predicted_probabilities,fill=room_type))+geom_histogram(binwidth = .1,alpha=.5)+theme_bw()+
  labs(x='Predicted Probabilities',y='Frequencies',title='Distribution of the predicted probabilities')

```




 
